|

### 2015-12-14 manuscript/manuscript_pve.Rmd
more small edits to manusscript....made in thesis word document too

##Reductions in growth and physiology under sink limitation
Soon after seedlings became established both height and diameter growth were negatively affected by decreasing soil volume. This led to the large reductions in biomass in small containers, compared to freely rooted seedlings. We analyzed the relationship between biomass growth and soil volume and found an increase of 34 % with a doubling of container volume, consistent with the meta-analysis of Poorter et al. [-@poorter2012pot]. These growth reductions were expected, as the impedance of root growth can cause reductions in overall plant growth and activity [@mcconnaughay1991physical; @young1997mechanical]. It has been shown that roots subjected to environmental stress may send inhibitory signals to the shoots that affect *g*~s~, cell expansion, cell division and the rate of leaf appearance [@passioura2002soil]. Here, this was evident in a large divergence in leaf area between seedlings in containers and free seedlings through time, with the eventual cessation of new leaf growth in seedlings in small containers.

@@

##Biomass partitioning under sink limitation
As biomass partitioning is likely controlled by the source and sink strength of all organs [@poorter2012biomass], it was important to determine which tissue components were most affected by the container size treatments. It was necessary to distinguish if growth was affected beyond ontogenetic constraints, by correcting for size, as biomass distribution is strongly size-dependent [@gould1966allometry; @lleonart2000removing]. In this study, there was no significant difference in root, leaf, or stem biomass partitioning with reduced soil volume compared to free seedlings, outside of ontogenetic drift (Figure `r allocfig`a,b). This is a surprising result as shifts in allocation have been noted specifically for nutrient limitation [@mcconnaughay1999biomass, and references therein]. Surprisingly, a constant ratio of fine root mass to leaf mass was observed across all treatments suggesting a functional partitioning response to optimize resource gain did not occur.

>>>>

##Reductions in growth and physiology under sink limitation
Soon after seedlings became established both height and diameter growth were negatively affected by decreasing soil volume. This led to the large reductions in biomass in small containers, compared to freely rooted seedlings. We analyzed the relationship between biomass growth and soil volume and found an increase of 34 % with a doubling of container volume, consistent with the meta-analysis of Poorter, Bühler, et al. [-@poorter2012pot]. These growth reductions were expected, as the impedance of root growth can cause reductions in overall plant growth and activity [@mcconnaughay1991physical; @young1997mechanical]. It has been shown that roots subjected to environmental stress may send inhibitory signals to the shoots that affect *g*~s~, cell expansion, cell division and the rate of leaf appearance [@passioura2002soil]. Here, this was evident in a large divergence in leaf area between seedlings in containers and free seedlings through time, with the eventual cessation of new leaf growth in seedlings in small containers.

@@

##Biomass partitioning under sink limitation
As biomass partitioning is likely controlled by the source and sink strength of all organs [@poorter2012biomass], it was important to determine which tissue components were most affected by the container size treatments. It was necessary to distinguish if growth was affected beyond ontogenetic constraints, by correcting for size, as biomass distribution is strongly size-dependent [@gould1966allometry; @lleonart2000removing]. In this study, there was no significant difference in root, leaf, or stem biomass partitioning with reduced soil volume compared to free seedlings, outside of ontogenetic drift (Figure `r allocfig`a,b). This is a surprising result as shifts in partitioning have been noted specifically for nutrient limitation [@mcconnaughay1999biomass, and references therein]. Surprisingly, a constant ratio of fine root mass to leaf mass was observed across all treatments suggesting a functional partitioning response to optimize resource gain did not occur.

### 2015-12-14 manuscript/thesis_pve.Rmd
more small edits to manusscript....made in thesis word document too

##2.5.1 Reductions in growth and physiology under sink limitation
Soon after seedlings became established both height and diameter growth were negatively affected by decreasing soil volume. This led to the large reductions in biomass in small containers, compared to freely rooted seedlings. We analyzed the relationship between biomass growth and soil volume and found an increase of 34 % with a doubling of container volume, consistent with the meta-analysis of Poorter et al. [-@poorter2012pot]. These growth reductions were expected, as the impedance of root growth can cause reductions in overall plant growth and activity [@mcconnaughay1991physical; @young1997mechanical]. It has been shown that roots subjected to environmental stress may send inhibitory signals to the shoots that affect g~s~, cell expansion, cell division and the rate of leaf appearance [@passioura2002soil]. Here, this was evident in a large divergence in leaf area between seedlings in containers and free seedlings through time, with the eventual cessation of new leaf growth in seedlings in small containers.

@@

##2.5.2 Biomass partitioning under sink limitation
As biomass partitioning is likely controlled by the source and sink strength of all organs [@poorter2012biomass], it was important to determine which tissue components were most affected by the container size treatments. It was necessary to distinguish if growth was affected beyond ontogenetic constraints, by correcting for size, as biomass distribution is strongly size-dependent [@gould1966allometry; @lleonart2000removing]. In this study, there was no significant difference in root, leaf, or stem biomass partitioning with reduced soil volume compared to free seedlings, outside of ontogenetic drift (Figure `r allocfig`a,b). This is a surprising result as shifts in allocation have been noted specifically for nutrient limitation [@mcconnaughay1999biomass, and references therein]. Surprisingly, a constant ratio of fine root mass to leaf mass was observed across all treatments suggesting a functional partitioning response to optimize resource gain did not occur.

>>>>

##2.5.1 Reductions in growth and physiology under sink limitation
Soon after seedlings became established both height and diameter growth were negatively affected by decreasing soil volume. This led to the large reductions in biomass in small containers, compared to freely rooted seedlings. We analyzed the relationship between biomass growth and soil volume and found an increase of 34 % with a doubling of container volume, consistent with the meta-analysis of Poorter, Bühler, et al. [-@poorter2012pot]. These growth reductions were expected, as the impedance of root growth can cause reductions in overall plant growth and activity [@mcconnaughay1991physical; @young1997mechanical]. It has been shown that roots subjected to environmental stress may send inhibitory signals to the shoots that affect g~s~, cell expansion, cell division and the rate of leaf appearance [@passioura2002soil]. Here, this was evident in a large divergence in leaf area between seedlings in containers and free seedlings through time, with the eventual cessation of new leaf growth in seedlings in small containers.

@@

##2.5.2 Biomass partitioning under sink limitation
As biomass partitioning is likely controlled by the source and sink strength of all organs [@poorter2012biomass], it was important to determine which tissue components were most affected by the container size treatments. It was necessary to distinguish if growth was affected beyond ontogenetic constraints, by correcting for size, as biomass distribution is strongly size-dependent [@gould1966allometry; @lleonart2000removing]. In this study, there was no significant difference in root, leaf, or stem biomass partitioning with reduced soil volume compared to free seedlings, outside of ontogenetic drift (Figure `r allocfig`a,b). This is a surprising result as shifts in partitioning have been noted specifically for nutrient limitation [@mcconnaughay1999biomass, and references therein]. Surprisingly, a constant ratio of fine root mass to leaf mass was observed across all treatments suggesting a functional partitioning response to optimize resource gain did not occur.
